Unsupervised Acquisition of Comprehensive Multiword Lexicons \\ using Competition in an $n$-gram Lattice

### 2017-05-25 reviewer_response.txt

> - I would like to see qualitative analysis of the sequences that were
identified as formulaic (or not) by annotators/the method. E.g.: How many of
them are syntactic constituents? Are there a wide variety of grammatical
kinds of expressions? Are high-frequency as well as low-frequency sequences
detected? What is the distribution in terms of length? What kinds of gapped
sequences are there? It would be nice to see some examples.

In the discussion, we have added a thorough analysis of specific examples and some statistics, including most of the information you ask for (implicitly if not explicitly).

> - Related to the previous point: While noting the distinction between FS and
MWEs, is it perhaps worth measuring recall against MWE datasets? Most MWEs
apart from proper names should count as FS, right?

Do you mean as a way of evaluating different systems built on the same corpus, or just to get a sense of coverage more generally? One central problem with using recall for evaluation is that in order to make it fair, there must be some control of the size of the lexicon. Neither our system nor our best performing competitors are ranking models, none of them offer an easy or fair way to simply take the best N words and look at recall. And even if they did, there's other kinds of biases that would come into play: for instance, PMI and similar metrics do well with rare terms that correspond well to the MWE lexicon of WordNet, for instance, which (at the type level) is dominated by certain kinds of terminology (biomedical terms, particularly) which are not really used outside a very specific community. This relates to a more fundamental problem with the idea of using recall in external lexicons as any kind of useful measure: we could not expect any particular corpus (even a very large one) to do justice to the full range of FS/MWE for a language, since most of which appear regularly only within specific genres. For example, we could not expect to build a good lexicon of English verb-particle constructions from a biomedical corpus, because many are too informal to be appropriate for that genre (e.g. "mess up"). With the qualitative analysis we have added, the genre-specificity of the FS we extra is hopefully more clear.

Though it perhaps does not directly address your comment (since it doesn't involve an external MWE lexicon), we were nonetheless inspired by your comment to include an analysis of the recall of MWE vs. non-MWE FS in the context of our existing test sets, and a small investigation of the recall of extremely common MWE from an existing MWE corpus, which sidesteps some of the genre bias issues mentioned above.

### 2017-05-25 tacl.tex
Added common MWE comparison, significance, reordered discussion

\begin{multline}
expl(t_1,...,t_N) = \prod_{i=1}^{N}{I_{on}(t_i)  C^{-\overlap(t_1)}} \\ + I_{\mathit{off}}(t_i) minLPR(t_i)^{-\cover(t_i) \cdot \clear(t_i)}
\end{multline}

When a node is off, its explainedness is the inverse of its minLPR, except if there are covering or clearing nodes which explain it by pushing the exponent of minLPR towards zero. When the node is on, its explainedness is the inverse of a fixed cost hyperparameter C, though this cost is increased if it overlaps with other active nodes. 
All else being equal, when $\minLPR(t) > C$, a node will be selected as an FS, and so, independent of the node interactions, C can be viewed as the threshold for the minLPR association measure under a traditional approach to MWE identification. There is no upper bound on $C$, but empirically, we have found values in the range $[3,6]$ give reasonable results for the languages presented in this paper.

@@

       \toprule
				& \multicolumn{7}{c}{\bf{English}} \\
       \cmidrule(lr){2-8}			
       & \multicolumn{3}{c}{\bf{ICWSM}} & &  \multicolumn{3}{c}{\bf{BNC}} & & \multicolumn{3}{c}{\bf{Croatian}}  && \multicolumn{3}{c}{\bf{Japanese}} \\
       \cmidrule(lr){2-4} \cmidrule(lr){6-8} \cmidrule(lr){10-12} \cmidrule(lr){14-16}
           \multicolumn{1}{c}{\bf{Source}}    & P & R & F &   & P & R & F &   & P & R & F &  & P & R & F \\
          \midrule 
Countrank & 0.13& 0.09 & 0.10 && 0.08 & 0.14 & 0.10 && 0.10 & 0.16 & 0.12 && 0.11 & 0.06 & 0.08 \\
PMIrank & 0.24& 0.15& 0.18 & & 0.12 & 0.25 & 0.16 & & 0.21 &0.32 & 0.26 & & 0.18& 0.08 & 0.11 \\ 
minLPRrank & 0.49& 0.31 & 0.39 & & 0.25& 0.44 & 0.32 & & 0.36 & 0.45 & 0.40 &  & 0.47 & 0.21 & 0.29 \\

@@

	Lattice\emph{$-$ovr} & 0.52& 0.51 & 0.51 & & 0.34& \bf{0.60} & 0.44& & 0.36 & \bf{0.67} & 0.47 & & 0.53 & \bf{0.62} & 0.58 \\

				Lattice & \bf{0.67} & 0.43 & \bf{0.52} & & \bf{0.40}& \bf{0.60} & \bf{0.48} & &\bf{0.44} & 0.56 & \bf{0.49} & & \bf{0.76} & 0.49 & \bf{0.59} \\ 
            \bottomrule

 \end{tabular}
  \caption{ Results of FS identification in various test sets; Countrank= lexicon created by ranking with frequency PMIrank = lexicon created with ranking pointwise mutual information, minLPRrank = lexicon created by ranking with minLPR, LPRseg = lexicon created by method of Brooke at \al (2015), -cl = no clearing, -ovr = no penalization of overlaps, P = Precision, R = Recall, F = F-score. Bold is best in column.}
	 \label{tab:main}

@@

The statistics in \tabref{tab:lexstats} otherwise reflect the quantity and diversity of FS across the corpora, particularly in terms of the number of POS patterns represented in the lexicon. Looking at the most common POS patterns across languages, only noun-noun and adjective-noun combinations ever account for more than 5\% of all word types in any of the lexicons. Though some of the diversity can of course be attributed to noise, it is safe to say that most FS do not fall into the standard two-word syntactic categories used in MWE work, and therefore identifying them requires a much more general approach like the one presented here.

@@

%The fact that less than 20\% of the FS derived from the BNC appear in the lexicon for the ICWSM corpus is fairly telling of how important genre differences are to the range of FS that will be seen in a particular corpus.
\tabref{tab:sample} contains 10 randomly selected examples from each of the corpora. Among the English examples, most of the clear errors are bigrams that reflect particular biases of their respective corpora: The phrase \ex{via slashdot} comes from boilerplate text identifying the source of a article, where as \ex{Maureen} (from \ex{Maureen says}) is a character in one of the novels included in the BNC. The longer FS mostly seem sensible, in that they are plausible lexicalized constructions, though \ex{primarily as \gap and} from the BNC is implausible as an FS and likely the result of noise. Some FS are dialectal variants, for instance \ex{license endorsed} refers to the British traffic violations.

What is perhaps most striking is the non-English examples is how poorly some FS translate: many good FS in these languages become extremely awkward when translated into English. This is expected of course for idioms like \ex{biti general poslije bitke} (be the general after the battle, \ie hindsight is 20/20), but it extends to other relatively compositional  constructions like \begin{CJK*}{UTF8}{min} \ex{こう\ 言う\ \gap\ が\ 続く} \end{CJK*} (repeat occurrences of \gap like this) and \ex{前期\ 比} (first half comparison). This highlights the importance of focusing on FS when learning a language, since many FS do not translate well (though some short ones do, particularly between the European languages). Though some of the errors seem to be the result of extra material added to an good FS, for instance \ex{promet teretnih vozila} (good vehicle traffic), most, again, are somewhat inexplicable artifacts of the corpus they were built from collected, like \ex{austrijski investitor} (Austrian investor).

Though we have demonstrated here that the idea of competition within a lattice is a powerful one for the production of FS lexicons, its usefulness derives less from the specific choices we have made in this instantation of the model, but rather the flexiblity that such a model provides for future research. Not only do alternatives like DP-seg and LocalMaxs fail to scale up, there are few obvious ways to improve on their simple underlying algorithms without destroying the elegance of the solution they provide. Fast and functional, the LPR decomp approach is nevertheless algorithmically ungainly, involving multiple layers of heuristic filtering. Our lattice method is aimed at something between these extremes: a practical, optimizable model, but with various component heuristics that can be improved upon. Though the current version of the clearing function is effective and has practical advantages relative to the other simple options that we tested, we think that function could be enhanced by more careful investigation of the statistical properties of FS, identifying to what extent the LPR of words around FS is likely to be inflated.

@@

%, and only the first two iterations require looking at a majority of the nodes, since there is no need to revisit a node if nothing has changed in its vicinity.

@@

Due to the lack of spaces between words and an agglutinative morphology, the standard approach to tokenization and lemmatization in Japanese involves morphological rather than word segmentation. In terms of the content of the resulting lexicon we believe the effect of this difference is modest, since much of what is extra in Japanese would simply be individual words in other languages (which could be viewed as being trivially part of the FS lexicon). However, from a theoretical perspective we might very much prefer to build FS for all languages starting from morphemes rather than words. Such a framework could, for instance, capture inflectional flexibilty versus fixedness directly in the model, with fixed inflectional morphemes included as an distinct element of the FS and flexible morphemes becoming gaps. However, for many languages this would result in a huge blow up in complexity with only modest increases in the scope of FS identification. Though it is indisputable that inflectional fixedness is part of the lexical information contained in an FS, in practice this sort of information can be efficiently derived post hoc using corpus statistics.

In \secref{sec:back}, we noted than FS is generally a much broader category than MWE, which we take as referring to terms which carry significant non-compositional meaning. We decided to investigate the distinction at a practical level by annotating the positive examples in the ICWSM test set for being MWE or non-MWE FS.\footnote{The set was exhaustively annotated by two annotators (&\kappa&: 0.73), and then conflicts were resolved through discussion.}. First, we note that only 0.28 of our FS types were labeled MWE; this is in contrast to, for instance, the annotation of \newcite{Schneider} where ``weak MWE'' make up a small fraction of MWE types. Even without any explicit representation of compositionality, our model did much better identifying MWE FS than non-MWE FS: recall for the former was 0.71, whereas recall for the latter was just 0.34. This may just reflect, however, the fact that a disproportionate number of MWE were noun-noun compounds, which are fairly easy for the model to identify.

Though the model presented here was designed specifically for FS extraction, we note that it could be useful for similar tasks such as unsupervised creation of morphological lexicons, particularly for agglutinative languages. Character or phoneme $n$-grams could compete in an identically structured lattice (though presumably without gapped $n$-grams) to be chosen as the best morphemes for the language, and we could straightforwardly adapt LPR to use phonological predictably (\ie based on vowel/consonant ``tags'') instead of syntactic predictability. It is likely, though, that further modifications would be necessary to target morphological phenomena well, and we leave this for future work.

>>>>

\begin{multline}
expl(t_1,...,t_N) = \prod_{i=1}^{N}{I_{on}(t_i)  C^{-\overlap(t_1)}} \\ + I_{\mathit{off}}(t_i) minLPR(t_i)^{-\cover(t_i) \cdot \clear(t_i)}
\end{multline}
\noindent
When a node is off, its explainedness is the inverse of its minLPR, except if there are covering or clearing nodes which explain it by pushing the exponent of minLPR towards zero. When the node is on, its explainedness is the inverse of a fixed cost hyperparameter C, though this cost is increased if it overlaps with other active nodes. 
All else being equal, when $\minLPR(t) > C$, a node will be selected as an FS, and so, independent of the node interactions, C can be viewed as the threshold for the minLPR association measure under a traditional approach to MWE identification. There is no upper bound on $C$, but empirically, we have found values in the range $[3,6]$ give reasonable results for the languages presented in this paper.

@@

       \toprule
				& \multicolumn{7}{c}{\bf{English}} \\
       \cmidrule(lr){2-8}			
       & \multicolumn{3}{c}{\bf{ICWSM}} & &  \multicolumn{3}{c}{\bf{BNC}} & & \multicolumn{3}{c}{\bf{Croatian}}  && \multicolumn{3}{c}{\bf{Japanese}} \\
       \cmidrule(lr){2-4} \cmidrule(lr){6-8} \cmidrule(lr){10-12} \cmidrule(lr){14-16}
           \multicolumn{1}{c}{\bf{Source}}    & P & R & F &   & P & R & F &   & P & R & F &  & P & R & F \\
          \midrule 
Countrank & 0.13& 0.09 & 0.10 && 0.08 & 0.14 & 0.10 && 0.10 & 0.16 & 0.12 && 0.11 & 0.06 & 0.08 \\
PMIrank & 0.24& 0.14& 0.18 & & 0.12 & 0.25 & 0.16 & & 0.21 &0.32 & 0.26 & & 0.18& 0.08 & 0.11 \\ 
minLPRrank & 0.49& 0.31 & 0.38 & & 0.25& 0.44 & 0.32 & & 0.36 & 0.45 & 0.40 &  & 0.47 & 0.21 & 0.29 \\

@@

	Lattice\emph{$-$ovr} & 0.52& 0.53 & 0.53 & & 0.34& \bf{0.60} & 0.44& & 0.36 & \bf{0.67} & 0.47 & & 0.53 & \bf{0.62} & 0.58 \\

				Lattice & \bf{0.67} & 0.44 & \bf{0.54} & & \bf{0.40}& \bf{0.60} & \bf{0.48} & &\bf{0.44} & 0.56 & \bf{0.49} & & \bf{0.76} & 0.49 & \bf{0.59} \\ 
            \bottomrule

 \end{tabular}
  \caption{ Results of FS identification in various test sets; Countrank= lexicon created by ranking with frequency PMIrank = lexicon created with ranking pointwise mutual information, minLPRrank = lexicon created by ranking with minLPR, LPRseg = lexicon created by method of Brooke at \al (2015), -cl = no clearing, -ovr = no penalization of overlaps, P = Precision, R = Recall, F = F-score. Bold is best in column. For the ICWSM only, difference relative to best baseline is significant at the $ p < 0.5$ level; the difference for all test sets considered together is significant at the $ p < 0.01$ level}
	 \label{tab:main}

@@

The statistics in \tabref{tab:lexstats} otherwise reflect the quantity and diversity of FS across the corpora, particularly in terms of the number of POS patterns represented in the lexicon. Looking at the most common POS patterns across languages, only noun-noun and adjective-noun combinations ever account for more than 5\% of all word types in any of the lexicons. Though some of the diversity can of course be attributed to noise, it is safe to say that most FS do not fall into the standard two-word syntactic categories used in MWE work, and therefore identifying them requires a much more general approach, like the one presented here.

@@

%The fact that less than 20\% of the FS derived from the BNC appear in the lexicon for the ICWSM corpus is fairly telling of how important genre differences are to the range of FS that will be seen in a particular corpus.
\tabref{tab:sample} contains 10 randomly selected examples from each of the corpora. Among the English examples, most of the clear errors are bigrams that reflect particular biases of their respective corpora: The phrase \ex{via slashdot} comes from boilerplate text identifying the source of a article, where as \ex{Maureen} (from \ex{Maureen says}) is a character in one of the novels included in the BNC. The longer FS mostly seem sensible, in that they are plausible lexicalized constructions, though \ex{primarily as \gap and} from the BNC is not obviously an FS and likely the result of noise. Some FS are dialectal variants, for instance \ex{license endorsed} refers to the British traffic violations. More generally, the FS lexicons created by these two corpora are quite distinct.

What is perhaps most striking is the non-English examples is how poorly some FS translate: many good FS in these languages become extremely awkward when translated into English. This is expected of course for idioms like \ex{biti general poslije bitke} (be the general after the battle, \ie hindsight is 20/20), but it extends to other relatively compositional  constructions like \begin{CJK*}{UTF8}{min} \ex{こう\ 言う\ \gap\ が\ 続く} \end{CJK*} (repeat occurrences of \gap like this) and \ex{前期\ 比} (first half comparison). This highlights the importance of focusing on FS when learning a language, since many FS do not translate well (though some short ones do, particularly for languages in the same family). Though some of the errors seem to be the result of extra material added to an good FS, for instance \ex{promet teretnih vozila} (good vehicle traffic), most, again, are somewhat inexplicable artifacts of the corpus they were built from collected, like \ex{austrijski investitor} (Austrian investor).

Since Zipfian frequency curves naturally extend to multiword FS vocabulary, our lexicons (and type-based evaluation of the them) are of course dominated by rarer terms. This is not, we would argue, a serious drawback, since in practical terms there is very little value in focusing on common MWEs like \ex{of course} which manually-built lexicons will cover; most of potential in automatic extraction comes from the long tail. However, we did investigate the other end of the Zipfian curve by extracting the 20 most common MWEs (including both strong and weak) annotated from the \newcite{Schneider14a} corpus. In the ICWSM lexicon, our recall for these common terms was fairly high (0.75), with errors seems to be mostly the result of longer phrases containing these terms ``winning out'' (in the lattice) over the shorter phrases, which have relatively low LPR due to extremely common constituent words; for example, we missed \ex{on time}, but had 19 FS which contain it, for instance \ex{right on time}, \ex{show up on time}, and \ex{start on time}. In one case which showed this same problem, \ex{waste * time}, the lexicon did have its ungapped version, highlighting potential for improvement.

In \secref{sec:background}, we noted than FS is generally a much broader category than MWE, which we take as referring to terms which carry significant non-compositional meaning. We decided to investigate the distinction at a practical level by annotating the positive examples in the ICWSM test set for being MWE or non-MWE FS.\footnote{The set was exhaustively annotated by two annotators ($\kappa$: 0.73), and then conflicts were resolved through discussion.}. First, we note that only 0.28 of our FS types were labeled MWE; this is in contrast to, for instance, the annotation of \newcite{Schneider14a} where ``weak'' MWE make up a small fraction of MWE types. Even without any explicit representation of compositionality, our model did much better identifying MWE FS than non-MWE FS: recall for the former was 0.71, whereas recall for the latter was just 0.34. This may simply reflect, however, the fact that a disproportionate number of MWE were noun-noun compounds, which are fairly easy for the model to identify.

Due to the lack of spaces between words and an agglutinative morphology, the standard approach to tokenization and lemmatization in Japanese involves morphological rather than word segmentation. In terms of the content of the resulting lexicon we believe the effect of this difference is modest, since much of the extra FS in Japanese would simply be individual words in other languages (which could be viewed as being trivially part of the FS lexicon). However, from a theoretical perspective we might very much prefer to build FS for all languages starting from morphemes rather than words. Such a framework could, for instance, capture inflectional flexibilty versus fixedness directly in the model, with fixed inflectional morphemes included as an distinct element of the FS and flexible morphemes becoming gaps. However, for many languages this would result in a huge blow up in complexity with only modest increases in the scope of FS identification. Though it is indisputable that inflectional fixedness is part of the lexical information contained in an FS, in practice this sort of information can be efficiently derived post hoc using corpus statistics.

Though we have demonstrated here that the idea of competition within a lattice is a powerful one for the production of FS lexicons, its usefulness derives less from the specific choices we have made in this instantation of the model, but rather the flexiblity that such a model provides for future research. Not only do alternatives like DP-seg and LocalMaxs fail to scale up to large corpora, there are few obvious ways to improve on their simple underlying algorithms without destroying the elegance of the solution they provide and worsening the tractability problem. Fast and functional, the LPR decomp approach is nevertheless algorithmically ungainly, involving multiple layers of heuristic-driven filtering with no possibility of correcting errors. Our lattice method is aimed at something between these extremes: a practical, optimizable model, but with various component heuristics that can be improved upon. For instance, though the current version of the clearing function is effective and has practical advantages relative to the other simple options that we tested during development, we think that function could be enhanced by more careful investigation of the statistical properties of FS, identifying to what extent the LPR of words around FS is likely to be inflated.

@@

@@

Finally, though the model presented here was designed specifically for FS extraction, we note that it could be useful for similar tasks such as unsupervised creation of morphological lexicons, particularly for agglutinative languages. Character or phoneme $n$-grams could compete in an identically structured lattice (though presumably without gapped $n$-grams) to be chosen as the best morphemes for the language, and we could straightforwardly adapt LPR to use phonological predictability (\ie based on vowel/consonant ``tags'') instead of syntactic predictability. It is likely, though, that further modifications would be necessary to target morphological phenomena well, and we leave this for future work.
